### 2021-06-14 README.Rmd
Added actions.

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Build Status](https://travis-ci.com/zabore/ppseq.svg?branch=main)](https://travis-ci.com/zabore/ppseq) [![CRAN status](https://www.r-pkg.org/badges/version/ppseq)](https://CRAN.R-project.org/package=ppseq)
[![Codecov test coverage](https://codecov.io/gh/zabore/ppseq/branch/main/graph/badge.svg)](https://codecov.io/gh/zabore/ppseq?branch=main)
[![R-CMD-check](https://github.com/zabore/ppseq/workflows/R-CMD-check/badge.svg)](https://github.com/zabore/ppseq/actions)
<!-- badges: end -->

The goal of ppseq is to provide functions that aid in designing clinical trials with sequential posterior probability monitoring incorporated.

## Installation

You can install the development version of ppseq from GitHub with:

```r
remotes::install_github("zabore/ppseq")
```