### 2022-03-28 README.Rmd
add cran installation to readme

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

<!-- badges: start -->
[![R-CMD-check](https://github.com/zabore/ppseq/workflows/R-CMD-check/badge.svg)](https://github.com/zabore/ppseq/actions)
[![Codecov test coverage](https://codecov.io/gh/zabore/ppseq/branch/main/graph/badge.svg)](https://codecov.io/gh/zabore/ppseq?branch=main)
<!-- badges: end -->

The goal of ppseq is to provide functions to design clinical trials using sequential predictive probability monitoring for futility stopping.

## Installation

You can install the production version of ppseq from CRAN with:

```r
install.packages("ppseq")
```

Or you can install the development version of ppseq from GitHub with:

```r
remotes::install_github("zabore/ppseq")
```