### 2022-08-03 README.Rmd
a few README updates

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->
[![R-CMD-check](https://github.com/zabore/ppseq/workflows/R-CMD-check/badge.svg)](https://github.com/zabore/ppseq/actions)
[![Codecov test coverage](https://codecov.io/gh/zabore/ppseq/branch/main/graph/badge.svg)](https://codecov.io/gh/zabore/ppseq?branch=main)
<!-- badges: end -->

<br>

## ppseq

The {ppseq} package provides functions to design clinical trials using Bayesian sequential predictive probability monitoring. Functionality is available to design [one-arm](https://www.emilyzabor.com/ppseq/articles/one_sample_expansion.html) or [two-arm](https://www.emilyzabor.com/ppseq/articles/two_sample_randomized.html) trials by searching over a grid of combinations of posterior and predictive thresholds and identifying the optimal design according to two criteria: accuracy and efficiency. Interactive plotting allows easy comparison of the various design options and easy trial implementation through decision rule plots.

## Installation

You can install the production version of ppseq from CRAN with:

```r
install.packages("ppseq")
```

Or you can install the development version of ppseq from GitHub with:

```r
remotes::install_github("zabore/ppseq")
```

## Basic usage